---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.19.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

## Association rules

Association rules find connections between variables which can be described as rules: if this, then that.
This and that can be single variables or combinations of variables, but they need to be binary (true/false).

```{r}
data <- read.csv("./data/daF3077_eng.csv", sep = ";")

summary( data )
```

```{r}
variables <- c("v1_1", "v1_2", "v1_3", "v1_4", "v1_5", "v1_6")

to_rules <- data[ , variables ] > 2

head( to_rules)
summary( to_rules )
```

## Finding the rules

With this command, potential rules are identified.

```{r}
library("arules")
rules <- apriori(to_rules,
    parameter = list(
        supp  = 0.001,
        conf  = 0.01,
        minlen = 3,
        maxlen = 10
    )
)
```

## Rules

Rules are of the form: `{v1_1} => {v1_2}`, which means that if `v1_1` is true, then `v1_2` is also likely to be true. 
The left-hand side of the rule is called the antecedent, and the right-hand side is called the consequent.

```{r}
inspect(rules)
```

# Tasks

* Give an interpretnation to the first rule in the output of `inspect(rules)`.
Share and discuss with a partner.


## Metrics: Lift, support, confidence

We were able to create a lot of rules.
But how do we know which ones are good?
We can use the following metrics:

* Support: How often does the rule apply to the dataset?
* Confidence: How often does the rhs happens when the lhs has happened?
* Lift: does lhs and rhs happen together more often than we would expect if they were independent?

Ideally you would have high support, high confidence and high lift.
Lift higher than 1 means its already better than random, but the higher the better.
Confidence and support vary from 0 to 1, the higher the better.

```{r}
rules <- apriori(to_rules,
    parameter = list(
        supp  = 0.01,
        conf  = 0.2,
        minlen = 1,
        maxlen = 10
    )
)

inspect(rules, by = "lift")
```

### Tasks

* What does it mean that the support is overall quite small? You can sort these by support by stating `by = "support"`.
* What is the highest confidence you can find and what does it mean, provide the rule an interpretation.
* Use some other variables to find rules which are relevant for social sciences.
Try to use the rules as a prediction machine, where there is an output variable you try to explain with other variables.


For R, also see [visualisation](https://github.com/mhahsler/arulesViz).

```{r}

```
