---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.19.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
library(caret)
library(kernlab)

data <- read.csv("./data/daF3077_eng.csv", sep = ";")
```

## Different algorithms

There are [various algorithms](https://topepo.github.io/caret/train-models-by-tag.html) available to conduct supervised machine learning.
As they work differently in statistical sense, some of them might work better for your prediction task.

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = data,
    method = "svmLinear",
)

summary(caret_model$finalModel)
```

### Tasks

* Add train-test split and evalute the model on the test set.
* Try out at least three different algorithms and compare their performance.
* Which of the algorithms you would use and why?


## Model tuning

In addition to various algorithms, each algorithm have parameters which impact how it operates.
They are in indicated in [the documentation](https://topepo.github.io/caret/train-models-by-tag.html).
Researchers systematically test different values out to examine what model is the best, i.e., lead to least errors.
This is called model tuning.

```{r}
intercept <- c(0, 1, 2, 3, 4)
```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = data,
    method = "lm",
    tuneGrid = data.frame(intercept = intercept)
)
```

```{r}
caret_model$results
```

#### Tasks

* Increase the number of values examined for intercept.
* Which intercept value would you choose?


## More parameters to tune

Above the linear model only had one paramter to tune, the intercept.
Other models can have more versatile set of parameters, leading one to establish of a grid of parameters to tune.


```{r}
library(nnet)
```

```{r}
parameters <- expand.grid(
    size = c(1, 2, 3, 4, 5),
    decay = c(0.1, 0.01, 0.001)
)

parameters
```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = data,
    method = "nnet",
    tuneGrid = parameters
)
```

```{r}
caret_model$results
```

#### Tasks

* Use a different model and different tuning parameters.
* Which model would you use


## Choosing the model

So we can fit several models on the data, with different parameters and different values.
How to choose the best one of them?
Using the `metric` function you can try to inform what is the `finalModel`

```{r}
caret_model$finalModel

```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = data,
    method = "nnet",
    tuneGrid = parameters,
    metric = "Rsquared"
)

caret_model$finalModel
```

#### Tasks

* Why do these two codes end up with different models?



