---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.19.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
data <- read.csv("./data/daF3077_eng.csv", sep = ";")
head( data )
```

## Traditional linear regression model

See variable descriptions in [here](https://services.fsd.tuni.fi/catalogue/FSD3077?tab=variables&study_language=en), but a quick summary:

* v4_1 I know which public welfare services and benefits are available in [country] for me and my family
* bv1 responder gender
* bv2 responder age
* bv4 ideology

Building a (poor) model to predict v4_1 from the background variables.

```{r}

model <- lm( v4_1 ~ bv1 + bv2 + bv4, data = data)
summary( model )
```

```{r}
table( data$v4_1 )
```

### Tasks

* Data contains also value 9 which stands for missing values. Remove it from the analysis.
* Add more data to the analysis


## The same with caret

[Caret](https://topepo.github.io/caret/index.html) is a R-library helping with supervised machine learning.
The results ought to be about the same.
However, we do a bit different evaluations on the model, instead of just $R^2$ we also examine mean-square error (as v4_1 is continous).

```{r}
library(caret)
library(ggplot2)
```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = data,
    method = "lm",
)

summary(caret_model$finalModel)
```

## Model fit

There are various ways to assess models and understand how well they fit to the data, as predictive models often make errors in the analysis.
One can calculate what is the difference between predicted values by the model and the actual (or: observed) values.
Various metrics follow this kind of thinking, including

* R2
* root mean square error (RMSE)
* mean absolute error (MAE)
* area under the curve (AUC)

Different research communities have somewhat different approaches on which of these to report.

```{r}
predicted_values <- predict(caret_model, newdata = data) ## this predicts using the model we just trained values based on the data.
postResample(pred = predicted_values, obs = data$v4_1)

ggplot( data = data.frame( observed = data$v4_1, predicted = predicted_values), aes(x = observed, y = predicted) ) +
    geom_point() +
    theme_minimal() +
    labs( title = "Observed vs Predicted Values", x = "Observed Values", y = "Predicted Values" )
```

### Tasks

* What does the figure above tell you about the model fit? Explain to a partner.
* Draw a plot showing the error and absolute errors per observe values


## Train-test split

Instead of using all materials for model training, let use only part of the data and then assess our performance with data not yet seen by the model.
This avoids overfit.

```{r}
trainIndex <- createDataPartition(data$v4_1, p = .8, 
                                  list = FALSE, 
                                  times = 1)

dataTrain <- data[ trainIndex,]
dataTest  <- data[-trainIndex,]

## check what the data look like

table( dataTrain$v4_1 )
table( dataTest$v4_1 )
```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = dataTrain,
    method = "lm",
)

summary(caret_model$finalModel)
```

```{r}
predicted_values <- predict(caret_model, newdata = dataTest) ## this predicts using the model with unused dataTest
postResample(pred = predicted_values, obs = dataTest$v4_1)
```

### Tasks

* Try different train-test split ratios
* How much does the performance drop between train data and test data? What does that tell you?


## Cross validation

```{r}
control <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated ten times
                           repeats = 10)
```

```{r}
caret_model <- train(
    v4_1 ~ bv1 + bv2 + bv4,
    data = dataTrain,
    method = "lm",
    trControl = control
)

summary(caret_model$finalModel)
```

```{r}
predicted_values <- predict(caret_model, newdata = dataTest) ## this predicts using the model with unused dataTest
postResample(pred = predicted_values, obs = dataTest$v4_1)
```

### Tasks

* Try different cross-validation rations?
* How much does the performance increase compared with non-cross-validated model?
