---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.19.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
data <- read.csv("./data/daF3077_eng.csv", sep = ";")
head( data )

## do you need to do cleaning of the data?
```

## Unsupervised learning

Unsupervised learning focus on working without a target variable.
The goal is to find patterns in the data, such as clusters of similar observations or associations between variables.
This means there is no _correct_ outcome like in supervised machine learning, but rather various different outcomes.

We examine how to cluster different welfare needs (`v1_x`) to summarise what kind of groups emerge in the needs.

```{r}
variables <- c("v1_1", "v1_2", "v1_3", "v1_4", "v1_5", "v1_6")
```

```{r}
to_cluster <- data[ , variables ]
summary( to_cluster )
```

## Clustering

Clustering groups together similar observations from the data.


### $k$-means

$k$-means method to identifies _centroids_ that summarise the response patterns of participants within each cluster (and assign each responder into one of the clusters).
[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/kmeans)

```{r}
clusters <- kmeans(to_cluster, centers = 3)
table(clusters$cluster)
```

### Hierarcical clustering

Hierarcical clustering operates by separating the data into branches based on similarity.
[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/hclust)

```{r}
dist_matrix <- dist(to_cluster)
clusters <- hclust(dist_matrix, method = "ward.D2")
plot(clusters)
```

```{r}
clusters$clusters <- cutree(clusters, k = 3)
table( clusters$clusters )
```

```{r}
library(dbscan)

clusters <- dbscan( to_cluster, eps = 1, minPts = 50 )
table( clusters$cluster )
```

## Tasks

* Check what these different clustering algorithms produce and describe the clusters.
* Check similarities and differences between clustering outcomes across methods.
* Vary the number of clusters and seeds to examine the stability of the outcomes.

```{r}
scaled <- scale( to_cluster )

head( scaled)
```

## Tasks

* Try out variables which were not all on the same scale.
How does the results differ on scaled and non-scaled clusters?


# Dimension reduction


### Principal component analysis

Reduces many correlated variables into a smaller set of components capturing the most variance in the data.
[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/prcomp)

```{r}
components <- prcomp(to_cluster, scale = TRUE)
```

```{r}
components$rotation
biplot(components,choices = c(1,3) )
```

### UMAP

A nonlinear method that maps high-dimensional data into a low-dimensional space.
Use `n_components` to change the number of coomponents coming out. [Documentation](https://www.rdocumentation.org/packages/umap/versions/0.2.10.0/topics/umap)

```{r}
library(umap)
```

```{r}
components <- umap( to_cluster, scale = TRUE, n_components = 5 )
```

```{r}
head( components$layout)
plot( components$layout[,1], components$layout[,3] )
```

## Tasks

* Add to above visualisations different genders and ages from the data.
* Add to above visualistions clusters from above.
* Try to interprent different components?
* Can you find a set of variables where components are more interpretable? 
* Try using PCA results in a regression model as predictors. Do they perform better than original variables?



